Colonies - Compute Continuums across Platforms

### 2023-04-04 paper_2023/paper.tex
wip

To address the aforementioned challenges, the paper introduces a framework termed Colonies which is available as open-source\footnote{https://github.com/colonyos/colonies} and designed to facilitate execution of computational workloads across a diverse range of platforms. It can be used to implement grid-alike distributed virtual computer systems, and is founded upon a loosely-coupled microservice architecture that beaks down complex workflows into composable functions. With the use of an HTTP protocol, these composable functions can be implemented in any computer language and be executed by independent executors deployed across various systems, e.g. cloud, edge, devices, or even in web browsers. Colonies orchestrates the execution and by using a zereo-trust security protocol, a collection of distributed executors can function as a single cohesive unit, thereby establishing seamless compute continuums across multiple platforms.

In addition to a technical description of the Colonies framework, the paper also describes some potential use cases. The paper describe how Colonies can be leveraged to build a remote sensing platform on Kubernetes, serve as a building block for edge computing, implement a serverless FaaS (Function-as-a-Service), and how it can be integrated with HPC platforms. Finally, the paper presents a performance investigation, as well as scalability and robustness evaluation.

@@

Additionally, many compute clusters operate on-premises installations. Sometimes it is necessary to temporarily increase the capacity of on-prem clusters by combining resources from multiple providers, for example, adding cloud compute resources to handle peak loads or utilize HPC resources to quickly reprocess historical data. Developing hybrid workflows where some jobs run in the cloud and others run on HPC systems requires even more software development efforts and is beyond the scope of most users, preventing them from utilizing powerful hardware. Clearly, there is a need for a framework that can consolidate various platforms to simplify development and enable seamless execution across platforms.

This paper presents a framework called Colonies, specifically designed for implementing distributed virtual computer systems that facilitate integration with other systems. Colonies is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment, offering a modern approach to grid computing. The main objective is to create a tool where monolithic workflows can be broken down into independent and separated compute units that can be dynamically added or removed while executing workflows. These compute units can be implemented in any computer language and easily integrated with other systems, and then be deployed anywhere on the Internet. The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable AI systems.

\section{Related work}
Workflow management has been extensively studied in both academic and industrial settings with numerous approaches proposed to address the challenges in this field. For example, Apache Airflow \cite{apache_airflow} is a popular open-source workflow management system for handling data engineering pipelines. Like Colonies, Apache Airflow enables developers to create custom operators and executors that can be integrated with various systems. Additionally, Apache Airflow offers an HTTP API that makes it possible to develop software development kits (SDKs) in various programming languages. However, Apache Airflow does not realy on a queuing system. Instead, it must be integrated with a message broker, such as RabbitMQ \cite{rabbitmq} or Kafka \cite{apache_kafka}, to implement task queues, resulting in a more complex architecture than Colonies. Furthermore, Colonies is based on a distributed microservice architecture that makes it more suitable for DevOps software development. As Colonies is loosely coupled, executors can be implemented and dynamically deployed without reconfiguring the workflow engine, effectively turning it to a grid computing platform.

Argo \cite{argowf} is an open source container-native workflow engine for orchestrating parallel jobs on Kubernetes. It is can be used for running CI/CD pipelines or compute intensive machine learning or data processing tasks where each job runs as a container. In contrast, Colonies offers a more versatile approach, allowing jobs to be launched within a container. As launching new containers on Kubernetes can occasionally be time-consuming, Colonies can delivers higher throughput as the costs of starting new jobs are minimal. This is particular useful when launching large container images, or workloads (e.g. Julia scripts) or that takes relative long time to start.

@@

J. Represa et al. \cite{GarciaRepresa1740746} explore various challenges associated with developing microservice-based workflow management for industrial automation within the context of the Arrowhead project. The authors conclude that microservice-based workflow technologies is viable for industrial applications, particularly due to their inherent flexibility. The primary contribution of this paper is a comprehensive technical description of how to implement a distributed workflow engine based on microservice principles, extending beyond orchestrating microservices to execute automation tasks.

Grid computing is a distributed computing model that allows multiple computers, sometimes geographically distributed, to work together as a cohesive system to solve large-scale computational problems. It can be used to harness resources from numerous computers to solve complex tasks.

>>>>

To address the aforementioned challenges, the paper introduces an open-source framework called Colonies\footnote{https://github.com/colonyos/colonies}, which is designed to facilitate execution of computational workloads across a diverse range of platforms. Colonies enables the development of a \emph{grid-like distributed virtual computer}, referred to as a \emph{colony}, which can effortlessly integrate with any third-party application or other workflow management systems.

Built on a modular microservice architecture, Colonies effectively decomposes complex workflows into highly composable functions. With the use of an HTTP protocol, these composable functions can be implemented in any computer language, and be executed by independent executors deployed across various systems, e.g. cloud, edge, high-performance computing (HPC) environments, devices, or even web browsers. By using a zero-trust security protocol, a collection of distributed executors can function as a single cohesive unit, thereby establishing seamless compute continuums across multiple platforms.

In addition to a technical description of the Colonies framework, the paper also describes some potential use cases. The paper describe how Colonies can be leveraged to build a remote sensing platform on Kubernetes, serve as a building block for edge computing, implement a serverless FaaS (Function-as-a-Service), and how it can be integrated with the Slurm workload manager. Finally, the paper presents a performance investigation, as well as scalability and robustness evaluation.

@@

Additionally, many compute clusters operate on-premises installations. Sometimes it is necessary to temporarily increase the capacity of on-prem clusters by combining resources from multiple providers, for example, adding cloud compute resources to handle peak loads or utilize HPC resources to quickly reprocess historical data. Developing hybrid workflows where some jobs run in the cloud and others run on HPC systems requires even more software development efforts \cite{wf_challenges} and is beyond the scope of many users, preventing them from utilizing powerful hardware. Clearly, there is a need for a framework that can consolidate various workflow management platforms to simplify development and enable seamless execution across platforms.

This paper presents a framework called Colonies, specifically designed to implement a distributed virtual computer system. Colonies is built around a loosely-coupled microservices architecture that separates workflow definitions from implementation and deployment. The primary objective is to establish a platform where monolithic workflows can be decomposed into independent computational units that can easily integrate with other systems or applications. These compute units can be implemented in any computer language and be deployed anywhere on the Internet. The remainder of the paper describes the Colonies framework and how it can be used to create robust and scalable AI systems.

\section{Related work}
Workflow management has been extensively studied in both academic and industrial settings with numerous approaches \cite{service_wfs, schmitt2022workflow, GarciaRepresa1740746, Ouyang2010, NIKOLOV2021100440, workflow_in_bigdata} proposed to address the challenges in this field. Recently, Apache Airflow \cite{apache_airflow} has become a popular open-source workflow management system for handling data engineering pipelines. Like Colonies, Apache Airflow enables developers to create custom operators and executors that can be integrated with various systems. Additionally, Apache Airflow offers an HTTP API that makes it possible to develop software development kits (SDKs) in various programming languages. However, Apache Airflow does not realy on a queuing system. Instead, it must be integrated with a message broker, such as RabbitMQ \cite{rabbitmq} or Kafka \cite{apache_kafka}, to implement task queues, resulting in a more complex architecture than Colonies. Furthermore, Colonies is based on a distributed microservice architecture that makes it more suitable for DevOps software development. As Colonies is loosely coupled, executors can be implemented and dynamically deployed without reconfiguring the workflow engine, effectively turning it to a grid computing platform.

Argo \cite{argowf} is an open source container-native workflow engine for orchestrating parallel jobs on Kubernetes. It is can be used for running CI/CD pipelines or compute intensive machine learning or data processing tasks where each job runs as a container. In contrast, Colonies offers a more versatile approach, allowing jobs to be launched within an already started container. As launching new containers on Kubernetes can occasionally be time-consuming, Colonies can delivers higher throughput as the costs of starting new jobs are minimal. This is particular useful when launching large container images, or workloads (e.g. Julia scripts) or that takes relative long time to start.

@@

J. Represa et al. \cite{GarciaRepresa1740746} explore various challenges associated with developing microservice-based workflow management for industrial automation within the context of the Arrowhead project. The authors conclude that microservice-based workflow technologies is viable for industrial applications, particularly due to their inherent flexibility. The primary contribution of this paper is a comprehensive technical description of how to implement a distributed workflow engine based on microservice principles, extending beyond orchestrating microservices to execute automation tasks within one platform. The vision is to create a distributed architecuture, where microservices can reside anywhere on the Internet and still function as a cohesive unit to execute cross-platform workflows.

Grid computing is a distributed computing model that allows multiple computers, sometimes geographically distributed, to work together to solve large-scale computational problems. It can be used to harness compute resources from numerous computers to solve complex tasks. To the authors best knowledge, no one has ever used to grid platform as an integration platform to orchestrate AI workload across platforms. The main contribution of this paper is a technical description of the Colonies framework and a description of potential use cases.
